Centers and centralizers of higher enveloping algebras of Lie algebras

### 2020-09-28 draft.tex
harish chandra

\subsection{Reductive Lie algebras and a generalized Harish-Chandra homomorphism}

@@

The following result generalizes the Harish--Chandra isomorphism to the $\EE_n$ center of the  enveloping $\EE_n$ algebra.

\begin{prp}
For $\fg$ a reductive \owen{semisimple?} Lie algebra with Cartan subalgebra~$\fh$ and Weyl group~$W$, 
there is a quasi-isomorphism of $\EE_{n+1}$ algebras
\[
{\frak Z}_n(\UU_n \fg) \simeq \clies (\fg) \otimes \Sym(\fh[1-n])^W.
\]
\end{prp}

\owen{I think the usual Chevalley theorem only applies when $1-n$ is an even integer. I'm not sure if or what the theorem is when $1-n$ is odd.}

\brian{Is there an explicit description of the $\EE_{n+1}$-structure for the RHS?}
\brian{Is the RHS formal, can we replace by Lie algebra cohomology.}

\begin{proof}
By the Poincar\'{e} lemma for de Rham forms and compactly supported de Rham forms, one has a canonical quasi-isomorphism
\[
\bZ_n(\UU_n \fg) \xto{\simeq} \clies\left(\fg, \Sym(\fg[1-n]) \right) .
\]
Since $\fg$ is reductive, there is a quasi-isomorphism \brian{do we actually need semisimple here?}
\[
\clies(\fg, \Sym(\fg[1-n]) \simeq \clies (\fg) \otimes \Sym(\fg[1-n])^G 
\]
where $G$ is a compact connected Lie group whose Lie algebra is $\fg$. 
Further, by the Chevalley restriction theorem one has 
\[
\Sym(\fg[1-n])^G \cong \Sym \left(\fh[1-n]\right)^W .
\]
\end{proof}

>>>>

\subsection{A generalized Harish-Chandra isomorphism}

@@

Using our models, we obtain a generalization of the Harish--Chandra isomorphism to the $\EE_n$ center of the  enveloping $\EE_n$ algebra. 
The following result follows from a simple application of the Poincar\'{e} lemma for de Rham forms and compactly supported de Rham forms.

\begin{prp}
There is a canonical quasi-isomorphism of $\EE_{n+1}$-algebras
\[
\bZ_n(\UU_n \fg) \xto{\simeq} \clies\left(\fg, \Sym(\fg[1-n]) \right) .
\]
\end{prp}

In the case that $\fg$ is semi-simple, there is the following modest refinement of the result. 
To state it, we set up some notation. 
Since $\fg$ is reductive, the $\fg$-invariants of the exterior algebra $\Sym(\fg[1])^\fg = \left(\oplus_{k \geq 0} \wedge^k \fg[k]\right)^\fg$ is isomorphic to the Lie algebra homology $H^{\rm Lie}_\bu(\fg)$ \cite{Cartan}.
Similarly, if $2\ell+1$ is an odd integer, then the $\fg$ invariants of $\Sym(\fg[k])^\fg = \oplus_{k \geq 0} \wedge^k \fg[2 \ell k + k]$ is isomorphic to the regraded Lie algebra homology $\Tilde{H}_\bu^{\rm Lie} (\fg, \ell)$
which in degrees $-2\ell k - k$, $k \in \ZZ$ is $H_{k}^{\rm Lie}(\fg)$ and zero in all other degrees.

\begin{prp}
For $\fg$ a semi-simple Lie algebra with Cartan subalgebra~$\fh$ and Weyl group~$W$, we have the following:
\begin{itemize}
\item if $n$ is odd, then there is a quasi-isomorphism of $\EE_{n+1}$-algebras
\[
{\frak Z}_n(\UU_n \fg) \simeq \clies (\fg) \otimes \Sym(\fh[1-n])^W.
\]
\item if $n$ is even, then there is a quasi-isomorphism of $\EE_{n+1}$-algebras
\[
{\frak Z}_n(\UU_n \fg) \simeq \clies (\fg) \otimes \Tilde{H}_\bu^{\rm Lie} \left(\fg, \frac{n}{2} -1\right).
\]
\end{itemize}
\end{prp}
\begin{proof}
By the Poincar\'{e} lemma for de Rham forms and compactly supported de Rham forms, one has a canonical quasi-isomorphism
\[
\bZ_n(\UU_n \fg) \xto{\simeq} \clies\left(\fg, \Sym(\fg[1-n]) \right) .
\]
Since $\fg$ is semi-simple, there is a quasi-isomorphism
\[
\clies(\fg, \Sym(\fg[1-n]) \simeq \clies (\fg) \otimes \Sym(\fg[1-n])^G 
\]
where $G$ is a compact connected Lie group whose Lie algebra is $\fg$.

Now, when $n$ is odd, we apply the Chevalley restriction theorem to obtain an isomorphism of graded vector spaces
\[
\Sym(\fg[1-n])^G \cong \Sym \left(\fh[1-n]\right)^W .
\]

Similarly, when $n$ is even, we apply the discussion in the paragraph preceding the statement of this proposition 
\end{proof}
